---
title: "Harmony_hamstring"
author: "Jolet Mimpen"
date: '2022-04-28'
output: html_document
---


```{r, message=FALSE, cache=FALSE, warning=FALSE}
knitr::opts_chunk$set(cache=FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(Matrix)
library(ggplot2)
library(scales)
library(Seurat)
library(dplyr)
library(patchwork)
library(harmony)
```

#Load filtered Seurat objects
```{r Read in RDS}

Hamstring0 <- readRDS("Ham0_filtered.RDS")
Hamstring1 <- readRDS("Ham1_filtered.RDS")
Hamstring2 <- readRDS("Ham2_filtered.RDS")
Hamstring5 <- readRDS("Ham5_filtered.RDS")
```

#Add metadata to each sample
```{r}

Hamstring0@meta.data$orig.ident <- "HAM_OMB0787"
Hamstring1@meta.data$orig.ident <- "HAM_OMB1139"
Hamstring2@meta.data$orig.ident <- "HAM_OMB1140"
Hamstring5@meta.data$orig.ident <- "HAM_OMB0782"

Hamstring0@meta.data$sex <- "Male"
Hamstring1@meta.data$sex <- "Male"
Hamstring2@meta.data$sex <- "Female"
Hamstring5@meta.data$sex <- "Female"

Hamstring0@meta.data$age <- "39"
Hamstring1@meta.data$age <- "24"
Hamstring2@meta.data$age <- "26"
Hamstring5@meta.data$age <- "18"

Hamstring0@meta.data$surgery <- "ACLR"
Hamstring1@meta.data$surgery <- "ACLR"
Hamstring2@meta.data$surgery <- "ACLR"
Hamstring5@meta.data$surgery <- "ACLR"

Hamstring0@meta.data$batch <- "28JAN2021"
Hamstring1@meta.data$batch <- "16MAR2021"
Hamstring2@meta.data$batch <- "16MAR2021"
Hamstring5@meta.data$batch <- "16MAR2021"

Hamstring0@meta.data$side <- "Right"
Hamstring1@meta.data$side <- "Right"
Hamstring2@meta.data$side <- "Right"
Hamstring5@meta.data$side <- "Left"

```

#Merge datasets
```{r merge all datasets}
data.list <- mget(ls(pattern = "Hamstring")) #findanything in global environment with the given name/pattern
data.integrated <- Reduce(function(x, y){merge(x,y)}, data.list)
remove(data.list) #cleanup
```


```{r typical seurat pathway}
data.integrated2 <- data.integrated %>% 
                FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
                ScaleData(verbose=FALSE) %>% 
                RunPCA(npcs = 50, verbose=FALSE)
```

#Check distribution by sample
```{r}
options(repr.plot.height = 5, repr.plot.width = 20)
p1 <- DimPlot(object = data.integrated2, reduction = "pca", pt.size = .1, group.by = "sample")
p2 <- VlnPlot(object = data.integrated2, features = "PC_1", group.by = "sample", pt.size = .1)
p1 + p2
remove(p1, p2)
```

#Check distrubtion by sex
```{r}
options(repr.plot.height = 5, repr.plot.width = 20)
p1 <- DimPlot(object = data.integrated2, reduction = "pca", pt.size = .1, group.by = "sex")
p2 <- VlnPlot(object = data.integrated2, features = "PC_1", group.by = "sex", pt.size = .1)
p1 + p2
remove(p1, p2)
```

#Check distrubtion by affected side
```{r}
options(repr.plot.height = 5, repr.plot.width = 20)
p1 <- DimPlot(object = data.integrated2, reduction = "pca", pt.size = .1, group.by = "side")
p2 <- VlnPlot(object = data.integrated2, features = "PC_1", group.by = "side", pt.size = .1)
p1 + p2
remove(p1, p2)
```

#Check distrubtion by batch
```{r}
options(repr.plot.height = 5, repr.plot.width = 20)
p1 <- DimPlot(object = data.integrated2, reduction = "pca", pt.size = .1, group.by = "batch")
p2 <- VlnPlot(object = data.integrated2, features = "PC_1", group.by = "batch", pt.size = .1)
p1 + p2
remove(p1, p2)
```

#Run Harmony
```{r}
data.integrated3 <- RunHarmony(data.integrated2, group.by.vars = "sample", plot_convergence = TRUE)
```

#Plot Dimplot and Elbowplot to pick dims for clustering
```{r}
all.genes <- rownames(data.integrated3)
data.integrated3 <- ScaleData(data.integrated3, features = all.genes) #scale
remove(all.genes) #cleanup code

data.integrated3 <- RunPCA(data.integrated3, features = VariableFeatures(object = data.integrated3)) #run pca, linear dimensional reduction

plot5 <- DimPlot(data.integrated3, reduction = "pca")

plot6 <- ElbowPlot(data.integrated3, ndims = 50, reduction = "pca") #determine dimensionality, pick dims for clustering

plot5
plot6
```

#Choose dimensions and resolution of clustering
```{r}
dims.choice <- 50
data.integrated3 <- data.integrated3 %>% 
  RunUMAP(reduction = "harmony", dims = 1:dims.choice) %>% 
  RunTSNE(reduction = "harmony", dims = 1:dims.choice) %>% 
  FindNeighbors(reduction = "harmony", dims = 1:dims.choice) %>% 
  FindClusters(resolution = 0.2) %>% 
  identity()
```


#Dimplots comparing different groups
```{r}
options(repr.plot.height = 4, repr.plot.width = 8)
DimPlot(data.integrated.labelled, reduction = "umap", group.by = "sample", pt.size = .1, split.by = 'sample')
DimPlot(data.integrated3, reduction = "umap", group.by = "sample", pt.size = .1)

DimPlot(data.integrated3, reduction = "umap", group.by = "sex", pt.size = .1, split.by = 'sex')
DimPlot(data.integrated3, reduction = "umap", group.by = "sex", pt.size = .1)

DimPlot(data.integrated3, reduction = "umap", group.by = "side", pt.size = .1, split.by = 'side')
DimPlot(data.integrated3, reduction = "umap", group.by = "side", pt.size = .1)

DimPlot(data.integrated3, reduction = "umap", group.by = "batch", pt.size = .1, split.by = 'batch')
DimPlot(data.integrated3, reduction = "umap", group.by = "batch", pt.size = .1)
```

#Dimplot and featureplot
```{r}
options(repr.plot.height = 4, repr.plot.width = 6)
DimPlot(data.integrated3, reduction = "umap", label = TRUE, pt.size = 0.7)
DimPlot(data.integrated3, reduction = "umap", label = FALSE, pt.size = 0.7)
DimPlot(data.integrated3, reduction = "tsne", label = FALSE, pt.size = 0.7)
FeaturePlot(data.integrated3, features = "nFeature_RNA")
FeaturePlot(data.integrated3, features = "mitoRatio")
```

#Bargraph for counts per cluster
```{r}
meta.data <- data.integrated.labelled[[]]

# create random classifications for the sake of this example
#meta.data$condition <- sample(c('A', 'B', 'C'), nrow(meta.data), replace = TRUE)

counts <- group_by(meta.data, sample, seurat_clusters) %>% summarise(count = n())

ggplot(counts, aes(seurat_clusters, count, fill = sample)) + geom_bar(stat = 'identity')
remove(meta.data, counts) #cleanup
```

#DE for each cluster
```{r}
#nclusters <- nlevels(so.filt$seurat_clusters)
for (i in 0:11){
  assign(paste0("Cluster", i), FindMarkers(data.integrated3, ident.1 = i, min.pct = 0.25))
}
remove(i) #cleanup
```


#Load SYMBOL name list
```{r get list of all features, for future conversion}
hgnc_list <- read_csv("/Volumes/Seagate Exp/Single nuc data/annotation.csv")

hgnc_list <- hgnc_list %>%
    dplyr::filter(!grepl("LRG_gene", hgnc_list$gene_biotype))
```



#Add SYMBOL names to clusters
```{r}
whereitis <- match(rownames(newCluster0), hgnc_list$EnsName)
newCluster0$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster1), hgnc_list$EnsName)
Cluster1$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster2), hgnc_list$EnsName)
Cluster2$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster3), hgnc_list$EnsName)
Cluster3$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster4), hgnc_list$EnsName)
Cluster4$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster5), hgnc_list$EnsName)
Cluster5$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster6), hgnc_list$EnsName)
Cluster6$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster7), hgnc_list$EnsName)
Cluster7$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster8), hgnc_list$EnsName)
Cluster8$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster9), hgnc_list$EnsName)
Cluster9$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster10), hgnc_list$EnsName)
Cluster10$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup

whereitis <- match(rownames(Cluster11), hgnc_list$EnsName)
Cluster11$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup


```



#Save cluster files
```{r}
write.csv(Cluster0, "HAMint3000_Cluster0_20220722.csv")
write.csv(Cluster1, "HAMint3000_Cluster1_20220722.csv")
write.csv(Cluster2, "HAMint3000_Cluster2_20220722.csv")
write.csv(Cluster3, "HAMint3000_Cluster3_20220722.csv")
write.csv(Cluster4, "HAMint3000_Cluster4_20220722.csv")
write.csv(Cluster5, "HAMint3000_Cluster5_20220722.csv")
write.csv(Cluster6, "HAMint3000_Cluster6_20220722.csv")
write.csv(Cluster7, "HAMint3000_Cluster7_20220722.csv")
write.csv(Cluster8, "HAMint3000_Cluster8_20220722.csv")
write.csv(Cluster9, "HAMint3000_Cluster9_20220722.csv")
write.csv(Cluster10, "HAMint3000_Cluster10_20220722.csv")
write.csv(Cluster11, "HAMint3000_Cluster11_20220722.csv")


```


#General overview
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
geneName <- c("TRDN", "DES", "TNNT3", "MYH1", "COL1A2", "COL3A1", "DCN", "NEGR1", "FBLN1", "TNNT1", "ATP2A2", "MYH7", "PIEZO2", "COL12A1", "MKX", "COL22A1", "CPM", "SORBS2", "PECAM1", "PTPRB", "FLT1", "VWF", "PAX7", "CALCR", "GREM1", "NOTCH3", "PDGFRB", "MYO1B", "MMRN1", "PROX1", "KDR", "FLT4", "MYH11", "NTRK3", "ITGA8", "GPAM", "AQP7", "ADIPOQ", "PTPRC", "CD247", "CD69", "BLNK", "CD163", "MSR1")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) + ggtitle("General") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Muscle-like cells
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
geneName <- c("TRDN", "DES", "ACTN2", "TNNC2", "TNNI2", "TNNT3", "ACTN3", "HOMER1", "KCNQ5", "MYH1", "MYH3", "MYLK4", "MYBPC2",  "TNNC1", "TNNI1", "TNNT1", "ATP2A2", "CCR3", "LGR5", "MYH7", "MYH7B", "NEK10", "TECRL", "COL22A1", "ADAMTSL1", "CPM", "CSMD1", "FRAS1", "PZP", "RALYL", "SAMD5", "SORBS2", "SPAG16")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) + ggtitle("Muscle-like cells") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Fibroblasts
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
geneName <- c("COL1A1", "COL1A2", "COL3A1", "COL3A2",  "DCN", 
              "ABCA8", "COL15A1", "CD55", "DCLK1", "ELN", "FBN1", "FBLN1", "FBLN2", "FBLN5", "KCND2", "NEGR1", "NOVA1", "VIT",
              "CADM1", "COL11A1", "COL11A2", "COL12A1", "COL24A1", "COMP", "CPXM2", "FMOD", "MET", "MKX", "ITGA10", "PIEZO2", "THBS4", "THSD4")

whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) + ggtitle("Fibroblasts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```




#Immune cells
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names

#CD69 = associated with different immune cells
#CD2 = on T cells and NK cells
geneName <- c("PTPRC", "CD2", "CD247", "IL7R", "CD27", "CD53", "CD69", "CD83", "MS4A1", "BLNK", "CD14", "CD163", "CD163L1", "MSR1", "DOCK2")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Immune cells") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```



#Endothelial cells
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
#PECAM (CD31) = Endothelial cell marker
#VCAM = Endothelial cell marker
#FLT1 = VEGF-receptor 1
#KDR = VEGF-receptor 2
#CD34 = 
#CDH5 = VE-cadherin
#NOTCH3 = smooth muscle cell marker (mural cells)
#ACTA2 = expressed by vSMC but not pericytes
#long list: geneName <- c("PECAM1", "VCAM1", "MCAM", "ICAM1", "FLT1", "KDR", "FLT4", "CD34", "CDH5", "NOTCH3", "PDGFRA", "PDGFRB", "ACTA2", "LYVE1", "VEGFA", "VEGFB", "VEGFC", "VEGFD", "CD248", "VWF", "ANGPT1", "MYO1A", "MYO1B", "ADAMTS12", "EBF2", "COL5A1", "COL5A2", "NAV1", "NAV3", "PGF", "SPARC", "NOTCH4", "CD44", "ITM2A", "EVA1", "LEFTY1", "LEFTY2", "GDF1", "EPHB4", "ENG")
geneName <- c("PECAM1", "PTPRB", "VWF", "ADGRL4", "CD34", "CDH5", "FLT1", "NOTCH4", "TEK", "NOTCH3", "ADAMTS12", "EBF2", "MYO1B", "PDE1C", "PDGFRB", "EPHB4", "FLT4", "KDR", "MMRN1", "NRG3", "PKHD1L1", "PROX1", "FLNA", "IRAG1", "ITGA8", "LMOD1", "MYH11", "NTRK3",  "SYNPO2")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Endothelial and smooth muscle cells") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Myosins
```{r find multiple genes and plot}
#NSG2
#CTNNA2 = regulate cell-cell adhestion and differentiation in nervous system
#DCC = mediates axon attraction of neuronal growth cones in the developing nervous system upon ligand binding

geneName <- c("ACTA2", "ACTG2", "CNN1", "MB", "MYH2", "MYH9", "MYH10", "MYH11", "MYH14", "MYL1", "MYL3", "MYL4", "MYL6", "MYL2", "MYL5", "MYL7", "MYL9")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Myosins") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Cluster9
```{r}
geneName <- c("ACTA2", "ACTG2", "CNN1", "FLNA", "IRAG1", "KCNMB1", "MYH11", "MYL9", "MYLK", "TAGLN", "ACTC1", "ASB2", "CASQ2", "CSDC2", "DACT3", "DBNDD2", "DES", "DNAJB5", "FBXL22", "FLNC", "IL17B", "ITGA8", "ITIH3", "JPH2", "KCNMA1", "LDB3", "LMOD1", "MUSTN1", "MYOCD", "NCS1", "NKD1", "NTRK3", "PDLIM3", "PLN", "PPP1R14A", "RASL12", "SCUBE3", "SMTN", "SYNM", "SYNPO2", "TGFB1I1", "TPM1", "TPM2", "TPM3", "TPM4", "WFDC1")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Smooth Muscle Cells") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```


#Extracellular matrix
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
#CRLR
geneName <- c("COL1A1", "COL1A2", "COL3A1", "COL4A1", "COL5A1", "COL5A2", "COL5A3", "COL11A1", "COL11A2", "COL12A1", "COL14A1", "COL15A1", "ELN", "PRG4", "FBN1", "FBN2", "FBN3", "EMILIN1", "COMP", "DCN", "BGN", "SPOCK1", "SPOCK2", "SPOCK3", "AMBP", "FMOD", "LUM", "ACAN", "HSPG2", "TGFBR3", "AGRN", "NCAN", "BCAN", "VCAN", "FBLN1", "FBLN2", "FBLN3", "FBLN4", "FBLN5", "LAMA1", "LAMA2", "LAMA3", "LAMA4", "LAMA5")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Extracellular matrix") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```


#All collagens
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names
#CRLR
geneName <- c("COL1A1", "COL1A2", "COL2A1", "COL3A1", "COL4A1","COL4A2","COL4A3", "COL4A4", "COL4A5", "COL4A6", "COL5A1", "COL5A2", "COL5A3", "COL6A1", "COL6A2", "COL6A3", "COL6A5", "COL7A1", "COL8A1", "COL8A2", "COL9A1", "COL9A2", "COL9A3", "COL10A1",  "COL11A1", "COL11A2", "COL12A1", "COL13A1", "COL14A1", "COL15A1", "COL16A1", "COL17A1", "COL18A1", "COL19A1", "COL20A1", "COL21A1", "COL22A1", "COL23A1", "COL24A1", "COL25A1", "EMID2", "COL27A1", "COL28A1", "COL29A1")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Collagens") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Adipose
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names

geneName <- c("PDE3B", "GPAM", "TMEM132C", "PLIN1", "PLIN4", "AQP7", "ADIPOQ")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Adipocytes") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Satellite cells
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names

geneName <- c("PAX7", "CALCR", "CDH4", "CLCN5", "CTNND2", "GREM1")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Satellite cells") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```


#Kendal paper, Figure 2A
```{r find multiple genes and plot}
#multiple genes find ENS name and create data frame with both names

geneName <- c("COL1A1", "COL1A2", "SCX", "FBN1", "MFAP5", "VCAN", "EMILIN1", "TGFB1", "LTBP1", "LTBP2", "COL4A1", "POSTN", "COL5A1", "COL5A2", "COL11A1", "ACAN", "NES", "TNC", "TPM2", "TAGLN", "MYL9", "ACTA2", "RGS5", "ITGA7", "COL14A1", "COL6A1", "COL6A2", "COL6A3", "COL3A1", "BGN", "FBLN1", "DPT", "CXCL14", "LY6E", "TPPP3", "PDGFRA", "FN1", "DCN", "LUM", "ASPN", "FMOD", "PRELP", "COMP", "TNXB", "PRG4", "ABI3BP", "CILP", "CILP2", "THBS4", "MYOC", "VTN", "CLEC3B", "CLEC15A1", "COL2A1", "OGN", "TNMD")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)
remove(whereitis, EnsName) #cleanup

#use this to plot and change the title for the gene name
#plot4 <- DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red"))
plot4 <- DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red"), idents = c("1, 3"))

#plot, remove the xlabels of Ens and replace with gene names
plot4 + scale_x_discrete(breaks=geneName$EnsName, labels=geneName$geneName) +ggtitle("Kendal paper, Figure 2") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```

#Find specific marker
```{r}
geneName <- c("NTS")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
geneName <- data.frame(geneName, EnsName)

FeaturePlot(data.integrated.labelled, features = geneName$EnsName) + ggtitle(geneName$geneName)
VlnPlot(data.integrated.labelled, features = geneName$EnsName) + ggtitle(geneName$geneName)
DotPlot(data.integrated.labelled, features = geneName$EnsName, cols = c("blue", "red")) + ggtitle(geneName$geneName)

#FeaturePlot(data.integrated3, features = geneName$EnsName) + ggtitle(geneName$geneName)
#VlnPlot(data.integrated3, features = geneName$EnsName) + ggtitle(geneName$geneName)
#DotPlot(data.integrated3, features = geneName$EnsName, cols = c("blue", "red")) + ggtitle(geneName$geneName)

remove(whereitis, EnsName, geneName) #cleanup
```



#Make heatmap of selected genes, with the option of only selecting certain clusters
#Troponins
```{r}
library(pheatmap)
#old version: geneName <- c("DCN", "TPM3", "DES", "TTN", "TRDN", "NCAM1", "LAMA2", "NOS1", "COL22A1", "KCNQ5", "CD36", "BICC1", "EGFR", "BICD1", "DEPTOR", "TNNT3", "TNNT1", "RYR3", "MYL3", "DOCK10", "MYH7", "MEG8", "CSMD1", "SORBS2", "OPCML", "PELI1", "ADAMTSL1", "STAC", "CUX2", "MMP2", "GRID1", "EBF1", "MYOC", "BMP5", "NEGR1", "NOVA1", "TSHZ2", "DLC1", "FBLN1", "MYBPC1", "EMC10")

geneName <- c("TNNT1", "TNNT2", "TNNT3", "TNNI1", "TNNI2", "TNNI3", "TNNC1", "TNNC2")
whereitis <- match(geneName, hgnc_list$geneName)
EnsName <- hgnc_list$EnsName[whereitis]
genes <- data.frame(geneName, EnsName)

genes <- intersect(genes$EnsName, rownames(GetAssayData(data.integrated.labelled, slot = 'data')))
mat <- AverageExpression(data.integrated.labelled, features = genes, slot = 'data')

mat1 <- t(scale(t(mat$RNA)))

#select which clusters you want to show using the following function; if you want to show all clusters, use mat1
mat2 <- subset(mat1, select = c("TNNT3+ TRDN+ cells", "TNNT1+ TRDN+ cells", "COL22A1+ TRDN+ cells"))

paletteLength = 50
myColor <- colorRampPalette(c("navy", "white", "red"))(50)
myBreaks <- c(seq(min(mat1), 0, length.out=ceiling(paletteLength/2) + 1), seq(max(mat1)/paletteLength, max(mat1), length.out=floor(paletteLength/2)))

pheatmap(mat2, 
         labels_row = geneName,
         na_col = "black",
         color = myColor,
         breaks = myBreaks,
         display_numbers = TRUE)

pheatmap(mat2, 
         labels_row = geneName,
         na_col = "black",
         display_numbers = TRUE)

```





#Comparing specific clusters
```{r comparing specific clusters}
# find all markers distinguishing Tenocytes A vs Tenocytes B
C1vsC3.markers <- FindMarkers(data.integrated3, ident.1 = c(1), ident.2 = c(3), min.pct = 0.25)
whereitis <- match(rownames(C1vsC3.markers), hgnc_list$EnsName)
C1vsC3.markers$geneName <- hgnc_list$geneName[whereitis]
remove(whereitis) #cleanup
#write.csv(Fib4vsFib7, "Fib4vsFib7.csv")

```

#GO Analysis with gProfiler 2
```{r}
library(gprofiler2)
set.seed(500)
topcluster <- Cluster4 %>% 
  filter(avg_log2FC > 0.5) %>%
  top_n(n=100, wt = avg_log2FC)
gost.res <- gost(topcluster$geneName, organism = "hsapiens", correction_method = "fdr", sources = "REAC")
p <- gostplot(gost.res, interactive = T, capped = T)
p
#publish_gosttable(gost.res, highlight_terms = NULL, use_colors = TRUE, show_columns = c("source", "term_name", "term_size", "intersection_size"), filename = NULL)

ReactomeCluster4 <- gost.res$result
```

#Reactome analysis dotplot Clusters 1 and 3
```{r}
Cluster1Reactome$Dataset = "Cluster1"
Cluster3Reactome$Dataset = "Cluster3"

ReactomeFibroblasts = merge(Cluster1Reactome, Cluster3Reactome, all=TRUE)

p <- ggplot(ReactomeFibroblasts, aes(x = precision, y = fct_reorder(term_name, precision))) + 
  geom_point(aes(size = precision, color = p_value)) +
  theme_bw(base_size = 14) +
  scale_colour_gradient(limits=c(0, 0.05), low="red") +
  ylab(NULL) +
  ggtitle("Reactome Fibroblasts")
p + facet_grid(.~Dataset)
```

#Reactome analysis dotplot Clusters 0 and 4
```{r}
ReactomeCluster0$Dataset = "Cluster0"
ReactomeCluster2$Dataset = "Cluster2"

ReactomeFibroblasts = merge(ReactomeCluster0, ReactomeCluster2, all=TRUE)

p <- ggplot(ReactomeFibroblasts, aes(x = precision, y = fct_reorder(term_name, precision))) + 
  geom_point(aes(size = precision, color = p_value)) +
  theme_bw(base_size = 14) +
  scale_colour_gradient(limits=c(0, 0.05), low="red") +
  ylab(NULL) +
  ggtitle("Reactome TRDN+ cells")
p + facet_grid(.~Dataset)
```

#Add labels for each cluster
```{r}
data.integrated.labelled <- data.integrated3
new.cluster.ids <- c("TNNT3+ TRDN+ cells", "FBLN+ Fibroblasts", "TNNT1+ TRDN+ cells", "MKX+ Fibroblasts", "COL22A1+ TRDN+ cells", "Vascular endothelium", "Satellite cells", "Mural cells", "Lymphatic endothelium", "Smooth Muscle Cells", "Adipocytes", "Immune cells")
names(new.cluster.ids) <- levels(data.integrated.labelled)
data.integrated.labelled <- RenameIdents(data.integrated.labelled, new.cluster.ids)
DimPlot(data.integrated.labelled, reduction = "umap", repel = TRUE, label = TRUE, label.size = 3, pt.size = 0.8) + NoLegend()
remove(new.cluster.ids)
```


#Save RDS file
```{r save}
#Ham.integrated <- data.integrated
saveRDS(data.integrated3.integrated, file="HAMSTRING_April2022.RDS")

```


